---
title: "GSS challange More Data"
author: "Wilhelmus Elerich"
date: "10/4/2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```
```{r eval=FALSE}
mydata <-mydata %>%
mutate(across(where(is.character), as.factor))
```
# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

H_0 Attitude on capital punishment is independent of race

H_1 Attitude on capital punishment is dependent on race

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
Start by looking at the data in the master branch:

```{R}
dim(mydata)
names(mydata)
```
Here we have two categorical variables; "race", which has three values, and "cappun", which has two values. Thus, we should use the cat_cat helper file for this problem. For our graphical analysis, we should use barplots. For numerical analysis, we shall create a table. For inferential results, we shall use the chi-squared test and the fisher test.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata, type="percent")

```  

Judging by this percentage bar plot, a large majority of white people favor capital punishment, while a small majority of black people are opposed, and an even smaller majority of other people favor capital punishment.

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```
Here we can see that 65.10% of white people support capital punishment, while only 55.20 percent of the "other" category also supports capital punishment. In contrast, only 42.63% of black people support capital punishment. The largest difference in opinion is that between white people and black people.
## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
```{r}
fisherXY(mydata$race,mydata$cappun)
```
Judging from our P-values, we can reject the null hypothesis, and state that there is a statistical difference between the views of white people on capital punishment and the views of black people on capital punishment. Judging by the odds ratios, we  can determine  that white people are around 2 and half times as likely to support capital punishment as black people, white people are around one and half times as likely to support capital punishment as "other" people, and black people are just above half as likely to support capital punishment as "other" people.

## More data version


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
